Sulfur Dioxide at Water's Surface

### 2011-02-18 Background.tex
Rearranging the filenames and working on the h2o-orientation results

	The analyses performed in this study relate results to the location of the gas-liquid interface. An interface, however, is a dynamic surface that is neither flat nor stationary. Previous studies have used the technique of fitting a lineshape to the density profile of the water or solute to extract interfacial shape and location parameters.\cite{Shamay2010,Wick2006c,Chowdhary2006} Hyperbolic tangent functions have been used for fitting, and values for the ``Gibb's dividing surface'' location, and interfacial width are determined.\cite{Matsumoto1988} However, in long simulations the location of the interface will change, and the net motion of surface waters will alter the interfacial width at any given timestep. Thus, the averaged density profile fitting will capture average widths and locations, not instantaneous values. The interface can be characterized by density profiles, molecular orientational profiles, or interfacial intermolecular correlations. In this work we approach the problem by finding a reference location for each timestep by averaging the positions of the waters contained in the topmost monolayer. This provides us with a consistent reference point in the simulations to which we relate our analyses. This method is independent of the various measures of the interface, but still provides an intuitive point of reference.

\subsection{Analysis of Orientation}

	The simulated systems were analyzed to characterize the orientation of \wat~and \suldiox~in various environments above, within, and below the surface region. The orientation of the surface molecules as \suldiox~adsorbs and interacts with water is a property of any hydrate complex formation that takes places. Better understanding of the molecular orientational distributions will help elucidate the chemistry occurring during the adsorption process.

	The two molecules studied, \wat~and \suldiox, and similarly shaped with a C2$_v$ axis along their bisectors, and a molecular plane is defined by their three atoms. An body-fixed frame is defined for \wat~and \suldiox~as shown in figure \ref{fig:molecular-frame}. In each analysis a world-fixed reference axis is used that is the long axis of the system's periodic cell (normal to the plane of the water surface) unless otherwise noted. The orientational analyses presented herein focus on two angles used to define molecular orientation in space.

	A ``tilt'' angle, $\theta$, defines the angle formed between the bisector vector (the molecular z-axis, pointing from the central atom in the direction of the other two atoms) and the positive system reference axis. Thus the value of $\theta$ falls within a range of $[0,\pi]$. An angle of $\theta=0$ indicates a molecule with its bisector aligned with the reference axis, while $\theta=\pi$ results from an anti-aligned configuration.

	A second angle, $\phi$, defines the molecular ``twist''. $\phi$ is the angle formed between the vector normal to the plane of the molecule (molecular y-axis) and the system reference axis. The values of $\phi$ fall in the interval $[-\pi,\pi]$. Given certain values of $\theta$, $\phi$ provides additional information about whether the molecular orientation is ``flat'' to the surface (e.g. the plane of the molecule is aligned with the plane of the surface), or if it is perpendicular. %However, the symmetry of the \wat~and \suldiox~molecules create equivalence between the two H, or O, atoms (in the \wat~or\suldiox, respectively). $\phi=-\pi$ is equivalent to $\phi=\pi$, and both situations are equivalent to $\phi=0$. Because $\phi=-\phi$, the value is reported within the range of $[0,\frac \pi 2]$. 
	Figure \ref{fig:water-angles} shows the angle definitions relating the body-fixed axes to a system-fixed reference axis.

>>>>

	The results of analyses performed in this study are related to the location of the liquid water surface. A surface, however, is a dynamic boundary that is neither flat or stationary. Several previous studies have used the technique of fitting a lineshape to the density profile of the water or solute to extract interfacial shape and location parameters.\cite{Shamay2010,Wick2006c,Chowdhary2006} Hyperbolic tangent functions have been used to extract fitting parameters, and values for the ``Gibb's dividing surface'' location, and interfacial width have thus been determined.\cite{Matsumoto1988} However, in long simulations the location and shape of the interface will change, and the motion of surface waters alters the interfacial width at any given timestep. Thus, the averaged density profile fitting will capture average widths and locations, not instantaneous values. Similarly, the averaged values of location and width will obscure information about any drift or deformation the surface undergoes. In this work we approach the problem of surface loation by calculating a reference location for each timestep by averaging the long-axis positions of the waters contained in the topmost monolayer. This provides us with a consistent reference point in the simulations to which we relate our analyses. This method provides an intuitive point of reference, but does not increase the computational burden of the analyses.

  The location of the water surface was thus determined by the averaged positions of a set number of the topmost waters in the simulated aqueous slabs. The number of waters counted was decided by taking a few issues into account. First, a visual inspection of the top 30x30 \angs$^2$ cross-section of the water slab over several timesteps indicated between 65-75 waters established a full monolayer. Alternatively, assuming a spherical model of water with a radius of 2.4\angs (the typical length of a bulk-water hydrogen bond), two layers of hexagonally tight-packed spheres would yield a maximum of 52 molecules. Increasing the number of waters used in calculating the surface location diminishes the effects of the few waters that briefly rise above the surface into the gas phase, stabilizing both the surface position and thickness. Taking the close-packed model as a minimum number of waters fit into a flat surface, the authors decided to use 70 water molecules for calculating surface location during the simulation analyses.

  %Taking one of the simulations of the neat-water slabs, a series of calculations were performed to find the averaged positions and the standard deviations of varying numbers of the topmost waters. It was found that using 70 waters produced a standard deviation of the water positions corresponding to a width of 3.2\angs. as reported by Dang et al.when fitting the density profile with a hyperbolic tangent lineshape. To match the interfacial width of Dang using the standard deviation of our simulations, 70 water molecules would have to be included in the top monolayer. Thus, the calculations and analyses presented in this work incorporate 70 water molecules when calculating the surface location.

\subsection{Analysis of Molecular Orientation}

	The simulated systems were analyzed to characterize the orientation of \wat~and \suldiox~in various environments above, within, and below the aqueous surface region. In order to better understand the chemistry occurring during the \suldiox~adsorption process, the molecular orientation of both the \was~and \suldiox~must be understood.  The two molecules studied are similarly shaped with a C2$_v$ axis along their bisectors, and a molecular plane defined by three atoms. For analysis, a body-fixed frame is defined for \wat~and \suldiox~as shown in figure \ref{fig:molecular-frame}. In each analysis a space-fixed reference axis is used that corresponds to the long axis of the system's periodic cell (normal to the plane of the water surface) unless otherwise noted. The orientational analyses presented herein focus on two angles used to define molecular orientation in space.

	A ``tilt'' angle, $\theta$, defines the angle formed between the molecular bisector vector (the molecular z-axis, pointing from the central atom in the direction of the other two atoms) and the positive system reference axis. Thus the value of $\theta$ falls within a range of $[0,\pi]$. An angle of $\cos(\theta)=1$ indicates a molecule with its bisector aligned with the reference axis, while $\cos(\theta)=-1$ results from an anti-aligned configuration.

	A second angle, $\phi$, defines the molecular ``twist'' for certain configurations of the molecule. $\phi$ is the angle formed between the vector normal to the plane of the molecule (molecular y-axis) and the system reference axis. The values of $\phi$ fall in the interval $[-\pi,\pi]$. Given certain values of $\theta$, $\phi$ provides additional information about whether the molecular orientation is ``flat'' to the surface (e.g. the plane of the molecule is aligned with the plane of the surface), or if it is perpendicular. %However, the symmetry of the \wat~and \suldiox~molecules create equivalence between the two H, or O, atoms (in the \wat~or\suldiox, respectively). $\phi=-\pi$ is equivalent to $\phi=\pi$, and both situations are equivalent to $\phi=0$. Because $\phi=-\phi$, the value is reported within the range of $[0,\frac \pi 2]$. 
	Figure \ref{fig:water-angles} shows the angle definitions relating the body-fixed axes to a system-fixed reference axis.

### 2011-02-18 EquilibratedOrientation.tex
Rearranging the filenames and working on the h2o-orientation results

Angular analyses were performed to characterize the net molecular orientation of \wat~and \suldiox~molecules. Histograms were created to show the angles $\theta$ and $\phi$ (see figure \ref{fig:water-angles}) relative to a positions from the top surface of the aqueous slab. Both \wat~and \suldiox~orientations are considered for both the neat-\wat~and saturated systems.

@@

The results of the analyses for \wat~are shown in figure \ref{fig:water-orientation} for both of the surface equilibrated systems. In both systems an orientational preference is found at both the top and bottom slab surfaces. Because the bottom slab is inverted with respect to the top, it is expected that the orientational distributions will reflect the symmetry. The bisector tilt concentrates around $\cos(\theta)=0$ within the first few\angs of the surface, and then the distribution becomes isotropic further into the water bulk. As the tilt nears $\cos(\theta)=0$ the \wat~bisector lies within the plane of the surface indicating a water orientation either flat on the surface, or with some amount of ``twist'' sending the OH bonds in towards, or out of the bulk. The value of $\phi$ determines the ``twist'' in this case. Both systems show a peak in the distributions around $\cos(\phi)=1$ at the water surfaces. This results from an orientation of the water's y-axis (normal to the molecular plane) aligned perpendicular to the plane of the water surface.

>>>>

Molecular orientation analyses were performed to characterize the net molecular orientation of \wat~and \suldiox~molecules in the different locations near and within the water surface. Histograms were created to show the angles $\theta$ and $\phi$ relative to positions from the top surface of the aqueous slab. Both \wat~and \suldiox~orientations are considered for both the neat-\wat~and saturated systems.

@@

The results of the analyses for \wat~are shown in figure \ref{fig:water-orientation} for both the neat-water and saturated systems. In both systems an orientational preference is found at the slab surfaces where the water is in contact with the gas phase, or the surface \suldiox~molecules. %The bisector tilt concentrates around $\cos(\theta)=0$ within the first few\angs of the surface, and then the distribution becomes isotropic further into the water bulk. As the tilt nears $\cos(\theta)=0$ the \wat~bisector lies within the plane of the surface indicating a water orientation either flat on the surface, or with some amount of ``twist'' sending the OH bonds in towards, or out of the bulk. The value of $\phi$ determines the ``twist'' in this case. Both systems show a peak in the distributions around $\cos(\phi)=1$ at the water surfaces. This results from an orientation of the water's y-axis (normal to the molecular plane) aligned perpendicular to the plane of the water surface.

### 2011-02-18 Methods.tex
Rearranging the filenames and working on the h2o-orientation results

Molecular dynamics simulations were performed using the Amber 11 software suite.\cite{Case2010} In this study we incorporate polarizable models for the \wat~and \suldiox~molecules. Fully polarizable models have been used previously in studies on interfacial systems because they are known to more accurately reproduce interfacial structure and free energy profiles.\cite{Wick2007,Rivera2006,Dang1998} The \wat~model used is the POL3 model,\cite{Caldwell1995} and for \suldiox we used the model of Dang, et al. that places a single polarizable center on the sulfur atom.\cite{Baer2010}

All simulations began with an equilibrated cube of 900 \wat~molecules, with sides of length 30\angs. The long axis of each simulation cell (the axis normal to the water surface) was set to 120\angs. The simulations all employed periodic boundaries to create an ``infinite-slab'' geometry. After equilibrating the neat-\wat~slabs for 1 ns, two types of systems were created by introducing \suldiox.

One system involved the addition of a single \suldiox~molecule either within the bulk of the water slab, or above the slab surface. The single-\suldiox~system was then evolved for 2 ns to produce an equilibrated starting configuration. The second system introduced 22 \suldiox~molecules into the water slab in order to saturate it to a level coinciding with the Henry's law constant for \suldiox~in water. Additionally, 50 \suldiox~molecules were introduced into the gas phase outside of the water slab to simulate an added 50 atm of \suldiox~gas pressure. The saturated system was then evolved for 2 ns to produce a starting configuration for further simulations.

@@

Equilibrium simulations involved adding \suldiox~to a water slab and equilibrating as outlined above. One system had a single \suldiox~added to the center of the water box, representing a concentration of 0.06 M. The complementary high-concentration system with 22 \suldiox~molecule in the bulk represents a system of concentration 1.35 M, with an additional 50 \suldiox~in the gas phase to model 50 atm of pressure above the water surface. Both the low and high concentration systems were then evolved for a further 2.5 ns with atomic coordinates written every 100 fs.

@@

% need to add in the specifics about the SMD routine, force constants, calculations, etc.
A second set of simulations involved preparing an equilibrated water slab as in the surface equilibrated method above. However, for both the low and high concentration systems, a single \suldiox~was introduced 20\angs above the slab, with the sulfur atom tethered to its initial position. The system was evolved for 1 ns, taking coordinate snapshots every 20 ps to create 50 starting points for further simulations. Steered molecular dynamics (SMD) were then performed on the 50 system configurations (in both the low and high concentrations) to guide the \suldiox~down towards a tethered water 15\angs under the water slab surface by applying a small steering force.\cite{Isralewitz2001} The \suldiox~thus passed through the continuum of environments from gas phase to hydrate complex formation or interaction with surface-bound \suldiox, water surface adsorption, and absorption into the bulk of the \wat~slab.

Figure \ref{fig:starting-configurations} illustrates two sample starting configurations for the SMD simulations, showing both the neat-\wat~slab and the saturated-\wat~slab configurations before steering the \suldiox~into the water bulk.

>>>>

Molecular dynamics simulations were all performed using the Amber 11 software suite.\cite{Case2010} Polarizable models for the \wat~and \suldiox~molecules were used in the simulations, and have been used previously in studies on interfacial systems because they are known to more accurately reproduce interfacial structure and free energy profiles.\cite{Wick2007,Rivera2006,Dang1998} The \wat~model used is the POL3 model,\cite{Caldwell1995} and for \suldiox we used the model of Dang, et al. that places a single polarizable center on the sulfur atom.\cite{Baer2010}

All simulations began with an equilibrated cube of 900 \wat~molecules, with sides of length 30\angs. The long axis of each simulation cell (the axis normal to the water surface) was then lengthened to 120\angs, and the systems were further equilibrated for 10 ns. The simulations all employed periodic boundaries to create an ``infinite-slab'' geometry. After equilibrating the neat-\wat~slabs two types of systems were created by introducing \suldiox: a single-\suldiox~system, herein referred to as the ``neat-water'' system, and a saturated \suldiox~system.

One system involved the addition of a single \suldiox~molecule either within the bulk of the water slab (for equilibrium MD), or above the slab surface (in the SMD simulations). The single-\suldiox~system was then evolved for 2 ns to produce an equilibrated starting configuration. The second system introduced 22 \suldiox~molecules into the water slab in order to saturate it to a level coinciding with the Henry's law constant for \suldiox~in water. Additionally, 50 \suldiox~molecules were introduced into the gas phase outside of the water slab, on both surfaces, to simulate an added 50 atm of \suldiox~gas pressure. The saturated system was then evolved for 2 ns to produce a starting configuration for further saturated simulations.

@@

Equilibrium simulations involved adding \suldiox~to a water slab and equilibrating as outlined above. One system had a single \suldiox~added to the center of the water box, representing a concentration of 0.06 M. The complementary saturated system with 22 \suldiox~molecule in the bulk represents a system of concentration 1.35 M, with an additional 50 \suldiox~in the gas phase to model 50 atm of pressure above the water surface. Both the low and high concentration systems were then evolved for a further 10 ns using a timestep of 0.5fs, with atomic coordinates written every 100 fs.

@@

% need to add in the specifics about the SMD routine, force constants, calculations, etc.
A second set of simulations involved preparing an equilibrated water slab as in the surface equilibrated method above. However, for both the neat-water and saturated systems, a single \suldiox~was introduced 20\angs above the water slab surface, with the sulfur atom tethered to its initial position. The systems were evolved for 1 ns, taking coordinate snapshots every 20 ps to create 50 starting points for further simulations. Steered molecular dynamics (SMD) were then performed on the 50 system configurations (in both the low and high concentrations) to guide the \suldiox~down towards a tethered water 15\angs under the water slab surface by applying a small steering force.\cite{Isralewitz2001} The \suldiox~thus passed through the continuum of environments from gas phase to (neat- and saturated) water surface adsorption, and finally absorption into the bulk of the \wat~slab.

Figure \ref{fig:starting-configurations} illustrates two sample starting configurations for the SMD simulations, showing both the neat-\wat~slab and the saturated-\wat~slab configurations before steering the \suldiox~towards the water bulk.
